Supplementary material

### 2022-02-08 submission/manuscript.tex
almost final version after comments

  \begin{abstract}
  \textbf{Objective}: To compare different risk-based methods for
  optimal prediction of individualized treatment effects from RCTs.
  \textbf{Study Design and Setting}: We simulated RCT data using diverse
  assumptions for the average treatment effect, a baseline prognostic
  index of risk (PI), the shape of its interaction with treatment (none,
  linear, quadratic or non-monotonic) and the magnitude of
  treatment-related harms (none or constant independent of the PI). In
  each sample we predicted absolute benefit using: models with a
  constant relative treatment effect; stratification in quarters of the
  PI; models including a linear interaction of treatment with the PI;
  models including an interaction of treatment with a restricted cubic
  spline (RCS) transformation of the PI; an adaptive approach using
  Akaike's Information Criterion. We evaluated predictive performance
  using root mean squared error and measures of discrimination and
  calibration for benefit. \textbf{Results}: The linear-interaction
  model and the RCS-interaction model outperformed the constant
  treatment effect model in many simulation scenarios. The RCS-model was
  optimal when quadratic or non-monotonic deviations from a constant
  treatment effect were stronger, and when sample size was larger.
  Larger sample size also supported the adaptive approach. Illustrations
  in the GUSTO-I trial confirmed these findings. \textbf{Conclusion}: An
  interaction between baseline risk and treatment assignment should be
  considered to improve treatment effect predictions. Non-linear RCS
  interactions should be considered only in larger sample sizes.
  \end{abstract}
   \begin{keyword} treatment effect heterogeneity absolute
benefit prediction models\end{keyword}
 \end{frontmatter}

@@

A \emph{stratified HTE method} has been suggested as an alternative to
traditional subgroup analyses. Patients are stratified into
equally-sized risk strata---in this case based on risk quartiles.
Absolute treatment effects within risk strata are estimated by the
difference in event rate between patients in the control arm and
patients in the treated arm. We considered this approach as a reference,
expecting it to perform worse than the other candidates, as its
objective is to provide an illustration of HTE rather than to optimize
individualized benefit predictions.

@@

We demonstrate the different methods for individualizing treatment
benefits using data from 30,510 patients with acute myocardial
infarction (MI) included in the GUSTO-I trial. 10,348 patients were
randomized to tissue plasminogen activator (tPA) treatment and 20,162
were randomized to streptokinase. The outcome of interest was 30-day
mortality (total of 2,128 events), recorded for all patients.

@@

The linear interaction model outperformed all RCS methods in terms of
RMSE in scenarios with true constant relative treatment effect (OR =
0.8, N = 4,250 and AUC = 0.75), strong linear and even strong quadratic
deviations from a constant relative treatment effect (Figure
\ref{fig:rmsebase}; panels A-C). However, with non-monotonic deviations
from a constant relative treatment effect, the RMSE of the linear
interaction model increased substantially, especially in the presence of
treatment-related harms (Figure \ref{fig:rmsebase}; panel D). In these
scenarios, RCS-3 outperformed all other methods in terms of RMSE. As
might be expected the constant treatment effect approach had overall
best performance under true constant treatment effect settings. It was
sensitive to all considered deviations, resulting in increased RMSE.
Finally, the adaptive approach had comparable performance to the
best-performing method in each scenario. However, in comparison with the
best-performing approach, its RMSE was more variable in the scenarios
with linear and non-monotonic deviations, especially when also including
moderate or strong treatment-related harms. On closer inspection, we
found that this behavior was caused by wrongly selecting the constant
treatment effect model in a substantial proportion of the replications
(Supplement, Figure S1). This problematic behavior was less with larger
sample sizes (see below).

@@

Increasing the sample size to 17,000 favored RCS-3 the most, It achieved
lowest or close to lowest RMSE across all scenarios (Figure
\ref{fig:rmsesamplesize}). Especially in cases of strong quadratic and
non-monotonic deviations RCS-3 had lower RMSE (median 0.011 for strong
quadratic deviations and 0.010 for non-monotonic deviations with no
treatment-related harms) compared to the linear interaction approach
(median 0.013 and 0.014, respectively), regardless of the strength of
treatment-related harms. Due to the large sample size, the RMSE of the
adaptive approach was even more similar to the best-performing method,
and the constant relative treatment effect model was less often wrongly
selected (Supplement, Figure S2).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/rmsesamplesize-1} \caption{RMSE of the considered methods across 500 replications calculated in simulated samples of size 17,000 rather than 4,250 in Figure \ref{fig:rmsebase}}\label{fig:rmsesamplesize}
\end{figure}

When we increased the AUC of the true prediction model to 0.85 (OR = 0.8
and N = 4,250). RCS-3 had the lowest RMSE in the case of strong
quadratic or non-monotonic deviations and very comparable performance to
the -- optimal -- linear interaction model in the case of strong linear
deviations (median RMSE 0.016 for RCS-3 compared to 0.014 for the linear
interaction model). As observed in the base case scenario the adaptive
approach wrongly selected the constant treatment effect model (23\% and
25\% of the replications in the strong linear and non-monotonic
deviation scenarios without treatment-related harms, respectively),
leading to more variability of the RMSE (Supplement, Figure S3).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/rmseauc-1} \caption{RMSE of the considered methods across 500 replications calculated in a simulated samples 4,250. True prediction AUC of 0.85.}\label{fig:rmseauc}
\end{figure}

In comparison with the true approach, dsicrimination for benefit in the
scenario with a constant relative treatment effect was only slightly
lower for the linear interaction model, but substantially lower for the
non-linear RCS approaches (Figure \ref{fig:discrimination}; panel A).
With strong linear or quadratic deviations from a constant relative
treatment effect, all methods discriminated quite similarly (Figure
\ref{fig:discrimination}; panels B-C). In the scenario with
non-monotonic deviations, the constant effect model had much lower
discriminative ability compared to all other methods (median AUC of
0.4971 for the constant effects model, 0.5285 for the linear interaction
model and 0.5304 for the best-performing RCS-3; Figure
\ref{fig:discrimination}; panel D). The adaptive approach was unstable
in terms of discrimination for benefit, especially in the presence of
treatment-related harms. With increasing number of RCS knots, we
observed decreasing median values and increasing variability of the
c-for-benefit in all scenarios. When we increased the sample size to
17,000 we observed similar trends, however the performance of all
methods was more stable (Supplement, Figure S4). Finally, when we
increased the true prediction AUC to 0.85 the adaptive approach in the
case of non-monotonic deviations was,

again, quite unstable, especially with null or moderate
treatment-related harms (Supplement, Figure S5). In these scenarios, the
adaptive approach tended to select more often the inferior constant
treatment effect method (Supplement, Figure S3)

@@

In terms of calibration for benefit, the constant effects model
outperformed all other models in the scenario with true constant
treatment effects, but was miscalibrated for all deviation scenarios
(Figure \ref{fig:calibration}). The linear interaction model showed best
or close to best calibration across all scenarios and only showed worse
calibration compared to RCS-3 in case of non-monotonic deviations and
treatment-related harms (Figure \ref{fig:calibration}; panel D). The
adaptive approach was worse calibrated in scenarios with strong linear
and non-monotonic deviations compared to the linear interaction model
and RCS-3. When we increased sample size to 17,000 similar conclusions
on calibration for benefit could be drawn. As expected, all methods
displayed more stable calibration performance due to the larger number
of patients (Supplement, Figure S6). When we increased the true
prediction AUC to 0.85, the linear interaction model was worse
calibrated, on average, than RCS-3 in the case of strong quadratic
deviations from constant relative treatment effects (Supplement, Figure
S7).

The results from all individual scenarios can be explored online at
\url{https://arekkas.shinyapps.io/simulation_viewer/}.

@@

\hypertarget{empirical-illustration-1}{%
\subsection{Empirical illustration}\label{empirical-illustration-1}}

In line with previous analyses {[}13,14{]}, we fitted a logistic
regression model with 6 baseline covariates, i.e.~age, Killip class,
systolic blood pressure, heart rate, an indicator of previous MI, and
the location of MI, to predict 30-day mortality risk. A constant effect
of treatment was included in the model. When deriving risk predictions
for individuals we set the treatment indicator to 0. More information on
model development can be found in the supplement (Section XX).

We used the risk linear predictor to fit a constant treatment effect
model, a linear interaction model and a RCS-3 model individualizing
absolute benefit predictions. RCS-4 and RCS-5 models were excluded. In
our simulations these methods were always outperformed by the simpler
approaches and were often overfitted. Finally, an adaptive approach with
only the 3 candidate models was also applied.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/gusto-1} \caption{Individualized absolute benefit predictions based on baseline risk when using a constant treatment effect approach, a linear interaction approach and RCS smoothing using 3,4 and 5 knots. Risk stratified estimates of absolute benefit are presented within quartiles of baseline risk as reference.}\label{fig:gusto}
\end{figure}

@@

The adaptive approach had adequate performance, following closely on
average the performance of the ``true'' model in most scenarios. With
smaller sample sizes it tended to miss the treatment-risk interactions
and selected simpler models (Supplementary Table S7). This conservative
behavior resulted in increased RMSE variability in these scenarios,
especially in the case of true strong linear or non-monotonic deviations
from the base case scenario. Therefore, in the case of smaller sample
sizes the simpler linear interaction model may be a safer choice for
predicting absolute benefits, if a non-constant treatment effect is
suspected.

Risk-based approaches to predictive HTE estimate treatment benefit as a
function of baseline risk. A limitation of our study is that we assumed
treatment benefit to be a function of baseline risk in the majority of
the simulation scenarios. We also considered constant moderate and
strong treatment-related harms, applied on the absolute scale to expand
the range of scenarios in line with previous work {[}15{]}. In a limited
set of scenarios where we assumed the existence of true
treatment-covariate interactions, our conclusions remained unchanged.
Even though the average error rates increased for all the considered
methods, due to the miss-specification of the outcome model, the linear
interaction model had the lowest error rates. RCS-3 had very comparable
performance. The constant treatment effect model often gave biased
results, espcially in the presence of moderate or strong
treatment-related harms. All the results of these simulations can be
found in Supplement, Section XX. Future simulation studies could explore
the effect of more extensive deviations from risk-based treatment
effects.

>>>>

  \begin{abstract}
  \textbf{Objective}: To compare different risk-based methods for
  optimal prediction of individualized treatment effects from RCTs.
  \textbf{Study Design and Setting}: We simulated RCT data using diverse
  assumptions for the average treatment effect, a baseline prognostic
  index of risk (PI), the shape of its interaction with treatment (none,
  linear, quadratic or non-monotonic) and the magnitude of
  treatment-related harms (none or constant independent of the PI). In
  each sample we predicted absolute benefit using: models with a
  constant relative treatment effect; stratification in quarters of the
  PI; models including a linear interaction of treatment with the PI;
  models including an interaction of treatment with a restricted cubic
  spline (RCS) transformation of the PI; an adaptive approach using
  Akaike's Information Criterion. We evaluated predictive performance
  using root mean squared error and measures of discrimination and
  calibration for benefit. \textbf{Results}: The linear-interaction
  model and the RCS-interaction model outperformed the constant
  treatment effect model in many simulation scenarios. The RCS-model was
  optimal when quadratic or non-monotonic deviations from a constant
  treatment effect were stronger, and when sample size was larger.
  Larger sample size also supported the adaptive approach. Illustrations
  in the GUSTO-I trial confirmed these findings. \textbf{Conclusion}: An
  interaction between baseline risk and treatment assignment should be
  considered to improve treatment effect predictions. Non-linear
  interactions should be considered only in larger sample sizes.
  \end{abstract}
   \begin{keyword} treatment effect heterogeneity absolute
benefit prediction models\end{keyword}
 \end{frontmatter}

@@

A \emph{stratified HTE method} has been suggested as an alternative to
traditional subgroup analyses {[}3,4{]}. Patients are stratified into
equally-sized risk strata---in this case based on risk quartiles.
Absolute treatment effects within risk strata are estimated by the
difference in event rate between patients in the control arm and
patients in the treated arm. We considered this approach as a reference,
expecting it to perform worse than the other candidates, as its
objective is to provide an illustration of HTE rather than to optimize
individualized benefit predictions.

@@

For each scenario setting we performed 500 replications, within which
all the considered models were fitted. For comparing between models we
simulated a super-population of size 500,000 for each scenario. We
calculated RMSE and discrimination and calibration for benefit of the
models derived in each replication of our simulation settings within
this super-population. \#\# Empirical illustration

We demonstrated the different methods for individualizing treatment
benefits using data from 30,510 patients with acute myocardial
infarction (MI) included in the GUSTO-I trial. 10,348 patients were
randomized to tissue plasminogen activator (tPA) treatment and 20,162
were randomized to streptokinase. The outcome of interest was 30-day
mortality (total of 2,128 events), recorded for all patients.

In line with previous analyses {[}13,14{]}, we fitted a logistic
regression model with 6 baseline covariates, i.e.~age, Killip class,
systolic blood pressure, heart rate, an indicator of previous MI, and
the location of MI, to predict 30-day mortality risk. A constant effect
of treatment was included in the model. When deriving risk predictions
for individuals we set the treatment indicator to 0. More information on
model development can be found in the supplement (Section 6).

@@

The linear interaction model outperformed all RCS methods in terms of
RMSE in scenarios with true constant relative treatment effect (OR =
0.8, N = 4,250 and AUC = 0.75), strong linear and even strong quadratic
deviations from a constant relative treatment effect (Figure
\ref{fig:rmsebase}; panels A-C). However, with non-monotonic deviations
from a constant relative treatment effect, the RMSE of the linear
interaction model increased substantially, especially in the presence of
treatment-related harms (Figure \ref{fig:rmsebase}; panel D). In these
scenarios, RCS-3 outperformed all other methods in terms of RMSE. As
might be expected the constant treatment effect approach had overall
best performance under true constant treatment effect settings. It was
sensitive to all considered deviations, resulting in increased RMSE.
Finally, the adaptive approach had comparable performance to the
best-performing method in each scenario. However, in comparison with the
best-performing approach, its RMSE was more variable in the scenarios
with linear and non-monotonic deviations, especially when also including
moderate or strong treatment-related harms. On closer inspection, we
found that this behavior was caused by wrongly selecting the constant
treatment effect model in a substantial proportion of the replications
(Supplement, Figure S3). This problematic behavior was less with larger
sample sizes (see below).

@@

Increasing the sample size to 17,000 favored RCS-3 the most, It achieved
lowest or close to lowest RMSE across all scenarios (Figure
\ref{fig:rmsesamplesize}). Especially in cases of strong quadratic and
non-monotonic deviations RCS-3 had lower RMSE (median 0.011 for strong
quadratic deviations and 0.010 for non-monotonic deviations with no
treatment-related harms) compared to the linear interaction approach
(median 0.013 and 0.014, respectively), regardless of the strength of
treatment-related harms. Due to the large sample size, the RMSE of the
adaptive approach was even more similar to the best-performing method,
and the constant relative treatment effect model was less often wrongly
selected (Supplement, Figure S4).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/rmsesamplesize-1} \caption{RMSE of the considered methods across 500 replications calculated in simulated samples of size 17,000 rather than 4,250 in Figure \ref{fig:rmsebase}. RMSE was calculated on a super-population of size 500,000}\label{fig:rmsesamplesize}
\end{figure}

When we increased the AUC of the true prediction model to 0.85 (OR = 0.8
and N = 4,250). RCS-3 had the lowest RMSE in the case of strong
quadratic or non-monotonic deviations and very comparable performance to
the -- optimal -- linear interaction model in the case of strong linear
deviations (median RMSE 0.016 for RCS-3 compared to 0.014 for the linear
interaction model). As observed in the base case scenario the adaptive
approach wrongly selected the constant treatment effect model (23\% and
25\% of the replications in the strong linear and non-monotonic
deviation scenarios without treatment-related harms, respectively),
leading to more variability of the RMSE (Supplement, Figure S5).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/rmseauc-1} \caption{RMSE of the considered methods across 500 replications calculated in simulated samples 4,250. True prediction AUC of 0.85. RMSE was calculated on a super-population of size 500,000}\label{fig:rmseauc}
\end{figure}

When assuming a true constant relative treatment effect, discrimination
for benefit was only slightly lower for the linear interaction model,
but substantially lower for the non-linear RCS approaches (Figure
\ref{fig:discrimination}; panel A). With strong linear or quadratic
deviations from a constant relative treatment effect, all methods
discriminated quite similarly (Figure \ref{fig:discrimination}; panels
B-C). In the scenario with non-monotonic deviations, the constant effect
model had much lower discriminative ability compared to all other
methods (median AUC of 0.4971 for the constant effects model, 0.5285 for
the linear interaction model and 0.5304 for the best-performing RCS-3;
Figure \ref{fig:discrimination}; panel D). The adaptive approach was
unstable in terms of discrimination for benefit, especially in the
presence of treatment-related harms. With increasing number of RCS
knots, we observed decreasing median values and increasing variability
of the c-for-benefit in all scenarios. When we increased the sample size
to 17,000 we observed similar trends, however the performance of all
methods was more stable (Supplement, Figure S6). Finally, when we
increased the true prediction AUC to 0.85 the adaptive approach in the
case of non-monotonic deviations was, again, more conservative,
especially with null or moderate treatment-related harms (Supplement,
Figure S7).

@@

In terms of calibration for benefit, the constant effects model
outperformed all other models in the scenario with true constant
treatment effects, but was miscalibrated for all deviation scenarios
(Figure \ref{fig:calibration}). The linear interaction model showed best
or close to best calibration across all scenarios and only showed worse
calibration compared to RCS-3 in case of non-monotonic deviations and
treatment-related harms (Figure \ref{fig:calibration}; panel D). The
adaptive approach was worse calibrated in scenarios with strong linear
and non-monotonic deviations compared to the linear interaction model
and RCS-3. When we increased sample size to 17,000 similar conclusions
on calibration for benefit could be drawn. As expected, all methods
displayed more stable calibration performance due to the larger number
of patients (Supplement, Figure S8). When we increased the true
prediction AUC to 0.85, the linear interaction model was worse
calibrated, on average, than RCS-3 in the case of strong quadratic
deviations from constant relative treatment effects (Supplement, Figure
S9).

The results from all individual scenarios can be explored online at
\url{https://arekkas.shinyapps.io/simulation_viewer/}. Additionally, all
the code for the simulations can be found at
\url{https://github.com/rekkasa/arekkas_HteSimulation_XXXX_2021}

@@

We used the derived prognostic index to fit a constant treatment effect
model, a linear interaction model and a RCS-3 model individualizing
absolute benefit predictions. RCS-4 and RCS-5 models were excluded. In
our simulations these methods were always outperformed by the simpler
approaches and were often overfitted. Finally, an adaptive approach with
only the 3 candidate models was also applied.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/gusto-1} \caption{Individualized absolute benefit predictions based on baseline risk when using a constant treatment effect approach, a linear interaction approach and RCS smoothing using 3 knots. Risk stratified estimates of absolute benefit are presented within quartiles of baseline risk as reference.}\label{fig:gusto}
\end{figure}

@@

The adaptive approach had adequate performance, following closely on
average the performance of the ``true'' model in most scenarios. With
smaller sample sizes it tended to miss the treatment-risk interactions
and selected simpler models (Supplement Section 4). This conservative
behavior resulted in increased RMSE variability in these scenarios,
especially in the case of true strong linear or non-monotonic deviations
from the base case scenario. Therefore, in the case of smaller sample
sizes the simpler linear interaction model may be a safer choice for
predicting absolute benefits, if a non-constant treatment effect is
suspected.

Risk-based approaches to predictive HTE estimate treatment benefit as a
function of baseline risk. A limitation of our study is that we assumed
treatment benefit to be a function of baseline risk in the majority of
the simulation scenarios. We also considered constant moderate and
strong treatment-related harms, applied on the absolute scale to expand
the range of scenarios in line with previous work {[}15{]}. In a limited
set of scenarios where we assumed the existence of true
treatment-covariate interactions, our conclusions remained unchanged.
Even though the average error rates increased for all the considered
methods, due to the miss-specification of the outcome model, the linear
interaction model had the lowest error rates. RCS-3 had very comparable
performance. The constant treatment effect model often gave biased
results, espcially in the presence of moderate or strong
treatment-related harms. All the results of these simulations can be
found in Supplement, Section 6. Future simulation studies could explore
the effect of more extensive deviations from risk-based treatment
effects.

### 2022-02-08 submission/supplement.rmd
almost final version after comments

@@

\begin{equation}
E\{y\given x,t_x\} = P(y\given x, t_x) = g(\beta_0+\beta_1x_1+\dots+\beta_8x_8+\gamma t_x)
\end{equation}
(\#eq:risk)
Individualized predictions are derived setting $t_x=0$.

## Approaches to individualize benefit predictions

### Risk stratification
Derive a prediction model using the same approach as above and divide
the population in equally sized risk-based subgroups. Estimate
subgroup-specific absolute benefit from the observed absolute
differences. Subject-specific benefit predictions are made by
attributing to individuals their corresponding subgroup-specific
estimate.

### Constant treatment effect
Assuming a constant relative treatment effect, fit the adjusted model
in \@ref(eq:risk). Then, an estimate of absolute benefit can be
derived from$$\hat{f}_{\text{benefit}}(lp\given \bm{x},
\hat{\bm{\beta}}) = g(lp) - g(lp+\hat{\gamma}) $$

### Linear interaction
The assumption of constant relative treatment effect is relaxed
modeling a linear interaction of treatment with the risk linear
predictor: $$E\{y\given \bm{x}, t_x, \hat{\bm{\beta}}\} =
g\big(lp+(\gamma_0+\gamma_1lp)t_x\big)$$ We predict absolute benefit
from $$\hat{f}_{\text{benefit}}(lp\given\bm{x}, \hat{\bm{\beta}}) =
g(lp) - g\big(\gamma_0+(1+\gamma_1)lp\big)$$

### Restricted cubic splines
Finally, we drop the linearity assumption and predict absolute benefit using
smoothing with restricted cubic splines with 3, 4, and 5 knots. More
specifically, we fit the model:

$$P(y=1\given lp, t_x) = g\big(\beta_0+\beta_{t_x}t_x+f_{RCS}(lp) + f_{RCS}(lp)\times t_x\big)$$
where
$$f_{RCS}(x)=\alpha_0+\alpha_1h_1(x)+\alpha_2h_2(x)+\dots+\alpha_{k-1}h_{k-1}(x)$$ with
$$h_{j+1}(x)= (x-t_j)^3-(x-t_{k-1})_+\frac{t_k-t_j}{t_k-t_{k-1}}+(x-t_k)^3_+\frac{t_{k-1}-t_j}{t_k-t_{k-   1}}$$
and $t_1,\dots,t_k$ are the selected knots. We predict absolute benefit from $$\hat{f}_{\text{benefit}}(lp \given \bm{x},\hat{\bm{\beta}})=P(y=1\given lp, t_x=0)-P(y=1\given lp,t_x=1)$$.

@@

```{r tab2, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
asd <- read_csv(
  here("data/processed/analysisIdsInteractions.csv")
) %>%
  select(-"type")
kableExtra::kbl(
  asd,
  format = "latex",
  escape = FALSE,
  longtable   = TRUE,
  align       = "r",
  booktabs    = TRUE,
  caption     = "Scenario settings of the entire simulation study.",
  col.names   = c(
    "Scenario",
    "Base",
    "N",
    "AUC",
    "Treatment-related harm",
    paste0("b", 0:8),
    "g0", "g1", "g2", "g5", "g6",
    "Before harms",
    "After harms"
  ),
  linesep = c('', '', '', '\\addlinespace'),
  digits      = 2,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 5,
      "Baseline risk"         = 9,
      "True treatment effect" = 5,
      "Benefit"               = 2
      )
  ) %>%
  kableExtra::kable_styling(font_size = 6)
```

@@

@@

\begin{equation}
P(outcome=1|X=x) = \text{expit}(lp(x)),
(\#eq:gusto1)
\end{equation}
where 
\begin{equation}
\begin{aligned}
lp(x)=& \beta_0 + \beta_1 \text{age} + \beta_2 I(\text{Killip}=II) + \beta_3I(\text{Killip}=III) +\\
&\beta_4 I(\text{Killip}=IV) + \beta_5ceil(\text{sysbp}, 120) + \beta_6 \text{pulse}+\\
&\beta_7 max(\text{pulse - 50, 0}) + \beta_8 I(\text{pmi}=yes)+\\
&\beta_9 I(\text{miloc}=Anterior) + \beta_9 I(\text{miloc}=Other) +\\
&\gamma\times\text{treatment}
\end{aligned}
(\#eq:gusto2)
\end{equation}

>>>>

\newpage
# Plausible scenario settings

In this section we present specific scenarios from our simulation settings
in which evolution of benefit followed similar patterns to (REF!!). In this case
patients were stratified into risk quarters based on their true baseline
risk. Within each risk quarter we constructed boxplots of true benefit.

```{r, cache=TRUE, echo=FALSE, fig.cap="Simulation scenarios that closely follow trials. In this case, we see increasing absolute benefits with increasing baseline risk.", fig.show="hold", out.width = '50%'}
include_graphics(
  c(
    here("figures/scenario_251.png"),
    here("figures/scenario_324.png")
  )
)
```

```{r, cache=TRUE, echo=FALSE, fig.cap="Simulation scenarios that closely follow trials. In this case, we see increasing absolute benefits with increasing baseline risk.", fig.show="hold", out.width = '50%'}
include_graphics(
  c(
    here("figures/scenario_406.png"),
    here("figures/scenario_422.png")
  )
)
```

# Approaches to individualize benefit predictions

@@

\begin{equation}
P(y=1\given\bm{x}, t_x) = g(\bm{x}^t\bm{\beta} + \gamma t_x)
\end{equation}
(\#eq:risk)

The risk linear predictor is then 
$lp(\bm{x}; \bm{\beta}) = \bm{x}^t\bm{\beta}$. 
We derive baseline risk predictions for patients by setting $t_x=0$ in
\@ref(eq:risk). All methods for individualizing benefit predictions are 2-stage
methods, that start by fitting a model for predicting baseline risk. The
estimated linear predictor of this model is

\begin{equation*}
\hat{lp} = lp(\bm{x};\hat{\bm{\beta}}) = \bm{x}^t\hat{\bm{\beta}}
\end{equation*}

## Risk stratification
Derive a prediction model using the same approach as above and divide the
population in equally sized risk-based subgroups. Estimate subgroup-specific
absolute benefit from the observed absolute differences. Subject-specific
benefit predictions are made by attributing to individuals their corresponding
subgroup-specific estimate.

## Constant treatment effect
Assuming a constant relative treatment effect, fit the adjusted model in
\@ref(eq:risk). Then, an estimate of absolute benefit can be derived from
$$\tau(lp;  \hat{\bm{\beta}}, \hat{\gamma}) = g(\hat{lp}) - g(\hat{lp}+\hat{\gamma}) $$

## Linear interaction
The assumption of constant relative treatment effect is relaxed modeling a
linear interaction of treatment with the risk linear predictor:

$$P(y=1\given lp, t_x;\hat{\bm{\beta}}) = g(\delta_0 + \delta_1\hat{lp} +
\gamma_0t_x + \gamma_1\hat{lp}t_x)$$

Then the predicted absolute benefit is 
$$\tau(lp;\hat{\bm{\beta}},\bm{\gamma}, \bm{\delta}) = g(\delta_0 +
\delta_1\hat{lp}) - g\big((\delta_0+\gamma_0) + (\delta_1+\gamma_1)\hat{lp}\big)$$

## Restricted cubic splines
Finally, we drop the linearity assumption and predict absolute benefit using
smoothing with restricted cubic splines with $k=3, 4$ and $5$ knots. More
specifically, we fit the model:

$$P(y=1|lp, t_x;\hat{\bm{\beta}}, \bm{\delta}, f_{RCS}) = g\big(\delta_0 + \gamma
t_x +f_{RCS}(\hat{lp})\times t_x\big)$$
where
$$f_{RCS}(x)=\alpha_0+\alpha_1h_1(x)+\alpha_2h_2(x)+\dots+\alpha_{k-1}h_{k-1}(x)$$
with $h_1(x)=x$ and for $j=2,\dots,k-2$
$$h_{j+1}(x)= (x-t_j)^3-(x-t_{k-1})_+^3 \frac{t_k-t_j}{t_k-t_{k-1}}+(x-t_k)^3_+\frac{t_{k-1}-t_j}{t_k-t_{k-1}}$$
where 
$t_1,\dots,t_k$ are the selected knots. We predict absolute benefit from 
$$ \tau(lp; \hat{\bm{\beta}}, \delta_0, f_{RCS}) = P(y=1\given\hat{lp},
t_x=0) - P(y=1\given \hat{lp}, t_x=1) $$

@@

```{r tab2, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
asd <- read_csv(
  here("data/processed/analysisIdsInteractions.csv")
) %>%
  select(-"type")
kableExtra::kbl(
  asd,
  format = "latex",
  escape = FALSE,
  longtable = TRUE,
  align = "r",
  booktabs = TRUE,
  caption = "Scenario settings of the treatment-covariate interaction scenarios.",
  col.names = c(
    "Scenario",
    "Base",
    "N",
    "AUC",
    "Treatment-related harm",
    paste0("b", 0:8),
    "g0", "g1", "g2", "g5", "g6",
    "Before harms",
    "After harms"
  ),
  linesep = c('', '', '', '\\addlinespace'),
  digits      = 2,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 5,
      "Baseline risk"         = 9,
      "True treatment effect" = 5,
      "Benefit"               = 2
      )
  ) %>%
  kableExtra::kable_styling(font_size = 6)
```

@@

For predicting baseline risk of 30-day mortality we fitted a logistic regression
model with age, Killip class (*Killip*), systolic blood pressure (*sysbp*),
pulse rate (*pulse*), prior myocardial infarction (*pmi*), location of
myocardial infarction (*miloc*) and treatment as the covariates. Baseline
predictions were made setting treatment to 0.

@@

\begin{equation*}
P(outcome=1|X=x) = \text{expit}(lp(x)),
(\#eq:gusto1)
\end{equation*}
where 
\begin{equation*}
\begin{aligned}
lp(x)=& \beta_0 + \beta_1 \text{age} + \beta_2 I(\text{Killip}=II) + \beta_3I(\text{Killip}=III) +\\
&\beta_4 I(\text{Killip}=IV) + \beta_5min(\text{sysbp}, 120) + \beta_6 \text{pulse}+\\
&\beta_7 max(\text{pulse - 50, 0}) + \beta_8 I(\text{pmi}=yes)+\\
&\beta_9 I(\text{miloc}=Anterior) + \beta_9 I(\text{miloc}=Other) +\\
&\gamma\times\text{treatment}
\end{aligned}
(\#eq:gusto2)
\end{equation*}
